---
title: "ESM206 Lab 5"
author: "Anna Abelman"
date: "10/29/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Lab 5 Objectives

- Learn to parse dates with "lubridate'
- get counts of observations with count()
- Uncount() rows
- One and two-sample t-test
- Create a heatmap with geom_tile()

```{r, include = FALSE}
#Read in data and attach packages
library(tidyverse)
library(here)
library(janitor)

lobster_abundance <- read_csv(here::here("data", "lobster_abundance.csv"),
                              na = "-99999") %>% 
  janitor::clean_names()

```

Use tidyr::uncount() function to convert out lobster data from frequency format to case format

```{r}
#not in tidy format because each observation is not in its own row
lobster_tidy <- lobster_abundance %>% 
  tidyr::uncount(lobster_count)
#now in tidy format!
```

#### Exploratory Data Visualization

We're only goin to consider 'site' as our variable of interest

```{r}
ggplot(lobster_tidy, aes(x = site, y = size_mm)) +
  geom_jitter(aes(color = site),
              alpha = 0.5,
              width = 0.2)

#histogram
ggplot(lobster_tidy, aes(x = size_mm)) +
  geom_histogram(aes(fill = site)) +
  facet_wrap(~site, scales = "free")

ggplot(lobster_tidy, aes(sample = size_mm)) +
  geom_qq() +
  facet_wrap(~site)


```

####Convert the 'date column to class 'Date'

We'll use the 'lubridate' package to convert to Date format, and then to help us easily parse month and year

```{r}

lobster_date <- lobster_tidy %>% 
  mutate(
    date_new = lubridate::mdy(date)
  )
```

Now let's parse year and month using lubridate::month() and lubridate::year()

```{r}
#creating new columns that only contain the month and then the year
lobster_parse_date <- lobster_date %>% 
  mutate(
    obs_month = lubridate::month(date_new, label = TRUE),
    obs_year = lubridate::year(date_new)
  )

#using label = TRUE shows the abbrev. for the month
```

Now let's find counts of observed lobsters based on different hierarchical groupings:

First: let's count lobsters by year and month
```{r}
lobster_ym <- lobster_parse_date %>% 
  dplyr::count(obs_year, obs_month)

lobster_ym

lobster_y <- lobster_parse_date %>% 
  count(obs_year)

lobster_y

lobster_site <- lobster_parse_date %>% 
  count(site)

lobster_site
```

If we want to crete a summary table that contains statistics OTHER than counts by group, it's easier to use group_by() + n()

```{r}
lobster_summary <- lobster_parse_date %>% 
  group_by(site) %>% 
  summarize(
    mean_size = mean(size_mm, na.rm = TRUE),
    sd_size = sd(size_mm, na.rm = TRUE),
    lobster_number = n()
  )

lobster_summary

#if you want to do multiple summary statistics, use group_by() and them summary to get all column, while using count will ONLY use count function
```

#### Find Confidence Intervals

Use t.test() function to find confidence intervals (for one sample) and perform t-tests to compare means of two sample (covered in lectures week 6)

```{r}
ivee_lobster <- lobster_tidy %>% 
  filter(site == "IVEE") %>% 
  pull(size_mm)

#pull create a vector that contains a series of values

t.test(ivee_lobster)

```

####Two-Sample t-test to Compare Means

We want to ask: Is there a significant difference in lobster lengths at Naples and Mohawk reefs?

We've done our necessary exploratory analyses to determine that a 2-sample t-test for means comparison is appropriate. 

```{r}
#two ways to do this:

#first:
naples_sample <- lobster_tidy %>% 
  filter(site == "NAPL") %>% 
  pull(size_mm)

mohk_sample <- lobster_tidy %>% 
  filter(site == "MOHK") %>% 
  pull(size_mm)

mn_ttest <- t.test(mohk_sample, naples_sample)

mn_ttest
#negative is based on the direction and order the vectors are in
```

There is a significant difference in lobster lengths between Naples and Mohawk Reef (t(df) = statistic, p < 0.001, alpha = 0.05)

```{r}
#second way:
lobster_mn <- lobster_tidy %>% 
  filter(site %in% c("NAPL", "MOHK"))

mn_ttest2 <- t.test(size_mm ~ site, lobster_mn)

mn_ttest2

mn_ttest2$p.value
mn_ttest2$statistic

```

Here is the p-value: `r mn_ttest2$p.value`

Here is my t-statistic: `r mn_ttest2$statistic`

#### Now: A Heatmap

```{r}
lobster_ys <- lobster_parse_date %>% 
  count(obs_year, site)

ggplot(lobster_ys, aes(x = obs_year, y = site)) +
  geom_tile(aes(fill = n))
```






